---
title: "CART Model"
output:
  html_document:
    df_print: paged
---


# Set Up

## Libraries

```{r echo=FALSE}
# Basic packages
#library(plyr)
library(tidyverse)
library(ggplot2)
library(lubridate)

library(tictoc)
#devtools::install_github("collectivemedia/tictoc")

# Packages for data analysis

# library(caret)            # For correlation matrices
# library(corrplot)         # Used for correlation statistics
# library(MASS)             # Used for regression work

# Packages for mapping
library(sp)
library(sf)
library(raster)
library(leaflet)

# Formatting and Visualizations
library(viridis)          # Need for viridis color palette in word cloud
library(knitr)            # Need to manage text outputs in RMarkdown
library(formatR)          # To manage margins

```

## Importing Data

```{r, Importing-the-Data}
data_cols <- cols(
     `Inspection ID` = col_double(),
     `DBA Name` = col_character(),
     `AKA Name` = col_character(),
     `License #` = col_double(),
     `Facility Type` = col_character(),
     Risk = col_factor(),
     Address = col_character(),
     City = col_character(),
     State = col_character(),
     Zip = col_double(),
     `Inspection Date` = col_datetime(format = ""),
     `Inspection Type` = col_character(),
     Results = col_character(),
     Violations = col_character(),
     Latitude = col_double(),
     Longitude = col_double(),
     Location = col_character()
)
raw.data <- read_csv("food-inspections.csv", col_types = data_cols)
```

## Communities

Chicago has community Areas and Neighborhoods. community areas have demographic data associated with them, but do not necessarily form homogeneous regions.

Shape files from [here](https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Neighborhoods/bbvz-uum9).

```{r Importing-Maps}
nbhd <- read_sf("ChicagoNbhd/geo_export_8956960c-bd3c-4e92-a8b2-4748382c28a3.shp")  %>%
     sf::st_transform('+proj=longlat +datum=WGS84')

communities <- read_sf("ChicagoCommunities/geo_export_55b54106-a17e-4776-86c9-12db49043238.shp")  %>%
     sf::st_transform('+proj=longlat +datum=WGS84')
```

This creates a new df which has a GIS geometry column in place of Latitude and Longitude.

```{r Mapping-Inspections}
inspections <- raw.data  %>%
     filter(! (is.na(Latitude) | is.na(Longitude)))  %>%
     add_column(community = NA_character_, .after = "Address")  %>%
     rowid_to_column  %>%
     st_as_sf(
          coords = c("Longitude", "Latitude"), 
          crs= crs(nbhd))
```

## Demographics

The following is a table of demographic area based on community Area
For working with demographic fields, added a second table that has demographic information preceded by "POP_".  Master demographic table looks for a prefix

```{r}
demographic_names <- c(
     'Layer', 'Name', 'GEOID', 'Population', 'Longitude', 'Latitude', 'Male',
     'Female', 'White', 'Black', 'Asian', 'Hispanic', 'Native', 'TwoOrMore',
     'Infant', 'Senior', 'Adult', 'MiddleAge', 'Young', 'Children','Juvenile'
)

demographics_columns <- cols(
     Layer = col_character(),
     Name = col_character(),
     GEOID = col_character(),
     Population = col_double(),
     Longitude = col_character(),
     Latitude = col_character(),
     `Male` = col_double(),
     `Female` = col_double(),
     `White` = col_double(),
     `Black` = col_double(),
     `Asian` = col_double(),
     `Hispanic` = col_double(),
     `Native` = col_double(),
     `TwoOrMore` = col_double(),
     `Infant` = col_double(),
     `Senior` = col_double(),
     `Adult` = col_double(),
     `MiddleAge` = col_double(),
     `Young` = col_double(),
     `Children` = col_double(),
     `Juvenile` = col_double()
)

demographics.info <- read_csv("Demographics/CHADDemographics.csv", 
                              col_names = demographic_names, 
                              skip = 1, 
                              n_max = 2,
                              show_col_types = FALSE) %>%
     dplyr::select('Male':'Juvenile')

demographics <- read_csv("Demographics/CHADDemographics.csv",
                         col_names = demographic_names, 
                         col_types = demographics_columns,
                         skip = 5
)

# Renaming variables to ensure consistency across program paradigms.
demographic.columns.fixed <- str_c(rep("POPCT_", 15), str_to_upper(demographic_names[7:21])) %>%
     append(demographic_names[1:6], after=0)

demographics <- demographics %>%
     setNames(demographic.columns.fixed) %>%
     mutate(Name = str_to_upper(Name))%>%
     as_tibble()

```

The table `demographicInfo` provides a description of the columns of `demographics`

Here, the `communities` and `demographics` dataframe are merged to create a `communities.demographics` dataframe.

```{r}
communities <- communities  %>%
     mutate(community = replace(community, community == "OHARE", "O'HARE"))

demographics <- demographics  %>%
     mutate(commID = str_to_upper(Name))

communities.demographics <- merge(communities, 
                                  demographics, 
                                  by.x = "community", 
                                  by.y = "commID")
```

Turning off the following section.
```{r other_demographics, eval=FALSE, include=FALSE}
demo.traf.vac <- read.csv("Demographics/Chicago_traffic_and_Vacant.csv") %>%
     mutate(Name=str_to_upper(Name))

demo.food.act <- read.csv("Demographics/Chicago_food_and_activity.csv") %>% 
     mutate(Name=str_to_upper(Name))

demo.health <- read.csv("Demographics/Chicago_Health_Status.csv") %>% 
     mutate(Name=str_to_upper(Name))

demo.income <- read.csv("Demographics/Chicago_income.csv") %>% 
     mutate(Name=str_to_upper(Name))

demo.lang.birth <- read.csv("Demographics/Chicago_lang_and_birth.csv") %>%
     mutate(Name=str_to_upper(Name))

demo.morbidity <- read.csv("Demographics/Chicago_Morbidity_Rate.csv") %>%
     mutate(Name=str_to_upper(Name))

demo.mortality <- read.csv("Demographics/Chicago_mortality.csv") %>%
     mutate(Name=str_to_upper(Name))

demo.population <- read.csv("Demographics/Chicago_pop.csv") %>% 
     mutate(Name=str_to_upper(Name))

#starts_with does not take regex.  Use matches()...

#combine_demo is a function for combining all the demographic files
#pre: dgraph is a nested tibble containing data tables in a column "demo"
#pre: dgraph contains variable search strings for each table ina column "prefix"
#pre: demo[i]] contains standardized demographic information
#pre: demo[[i]] contains community names in a column, Names.
#pre: demo[[i]] contains demographic infromation in one or more columns
#     preceded by "prefix"

#design: extra columns in demographic are excluded
#design: pivot longer combines all columns into a long format
#design: each demographic is appended to data file

#post: Returns a tibble containing community name, demographic name, year of demographic, and data value

combine_demo<-function(dgraph){
     x<-tibble()
     for(i in 1:length(dgraph$prefix)){
          x<-dgraph$demo[[i]]%>%
               dplyr::select(!c(Layer, GEOID, Latitude, Longitude))%>%
               pivot_longer(cols=matches(dgraph$prefix[i]),           
                            names_to=c("demo", "year"),
                            names_sep = "\\_")%>%
               add_row(x)
          
     }
     return(x)
}

#List of demographic tibbles and prefixes used for data within.  
demo_list <- tibble(
     demo = list(
          demo.traf.vac, demo.food.act, demo.health, demo.income, demo.lang.birth, 
          demo.morbidity, demo.mortality, demo.population, demo.population
     ), 
     prefix = c("TRF|VAC", "HC", "HC","ED|PCI|POV","LEQ|FOR",
           "HC","VR","VR|POP","POP|Pop" )
     )

demo.all <- combine_demo(demo_list)
#levels(as.factor(demo.all$demo))

#wide version of the demographic table
#Needed for statistical analysis

# demo.all.wide <- demo.all %>%
#      pivot_wider(names_from = c(demo,year))

```

# Cleaning Data

## Labeling Violations with community

We look for the intersections of `communities` and `inspection` and create `vio.comm.int`, which is a sparse matrix. For a community Area described in row `i` of `communities` there is a row in `vio.comm.int` which lists the rows `j` in `inspections` which fall within its boundary.

```{r Intersecting-Inspections-Communities}
tic("Running st_intersects")
vio.comm.int <- st_intersects(communities, inspections)
toc()
```

Now, we revise `inspections` to include a column listing the community Area the location falls within.

```{r Label Inspection by Community Area}
tic("Label Inspections by Community Area")
inspections <- inspections  %>% as_tibble()
for (i in 1:nrow(communities)) {
     inspections <- inspections %>%
          rows_update(
               tibble(
                    rowid = unlist(vio.comm.int[[i]]), 
                    community = communities$community[i]
               ),
               by = "rowid"
          )
}
inspections <- inspections  %>% st_as_sf()
toc()
```

## Risk
Update the risk category to Low / Medium / High
```{r}
inspections <- inspections  %>%
     mutate(
          Risk = fct_recode(inspections$Risk,
                            Low = "Risk 3 (Low)", 
                            Medium = "Risk 2 (Medium)", 
                            High = "Risk 1 (High)")) 
```

## Facility Type

Cleaning up by Facility type.  Also, remove items after inspection paradigm changed (2018-07-01 and select only restaurants).

would it be worth condensing mobile food dispensers and preparers into a single mobile category...?
Eliminates 9 categories and condenses 1500 entries
I think there is a compelling story to look at mobile facilities together.
There can always be a sub-analysis as time permits.
check and see if this is wrigley field only based on geo location.  MRB 3/24/2022

Rationale:  If a facility contains a restaurant, treat as a restaurant.
Then combine items as grocery stores.
looking at the number of combination grocery stores (LT 100), I think it makes sense to combine with general grocery

```{r}
#For consistency, first convert facilities to lowercase.
#Then, limit to inspection dates before paradigm changed.
#Look for any odd cases where a facility contains restaurant in name or is misspelled.  
inspections <- inspections  %>%
  mutate(`Facility Type` = str_to_lower(`Facility Type`)) %>%   
  dplyr::filter(`Inspection Date`<"2018-07-01", 
                `Facility Type`=="restaurant")
```

## Violation Field

Collect subset of violations.  

```{r}

#pre: Inspection data is filtered to prior to 2018-07-01
#pre: Inspection is filtered to facility == restaurant.  
tic("Splitting Violations")

inspections <- inspections  %>% 
     mutate(
          `Violations` = if_else(is.na(`Violations`), 
                                 true =  "0. NO VIOLATION RECORDED - Comments: NO VIOLATION RECORDED", 
                                 false = `Violations`),
     )

violations <- inspections  %>%
     dplyr::select(c("Inspection ID", "Violations"))  %>%
     separate_rows(Violations, sep="\\|")  %>%
     separate(Violations, c("Violation", "Observation"), 
              sep = "- Comments: ",
              fill = "right") %>% # Fill makes empty comments without complaining
     separate(Violation, c("V_Num", "V_Desc"), 
              sep = "\\.\\s",
              extra="merge" # Only split on the first . 
              )
violations <- violations %>%
     mutate(
          V_Num = str_replace_all(violations$V_Num,"\\s",""),
          V_Desc = str_replace_all(V_Desc,"^\\s",""),
          V_Desc = str_replace_all(V_Desc,"\\s$",""),
          V_Num = as.integer(V_Num))
toc()
```

## Remove Geometry

Geometry is dropped from here on out to simplify issues.

```{r data_reshape}
#drop geometry for simpicity of inspection.  
violations <- st_drop_geometry(violations)
inspections <- st_drop_geometry(inspections)

#inspect_long is key combination for downstream calcs and figs
inspect_long <- inspections %>%
     dplyr::select(-c(Violations, Location))%>%
     inner_join(violations, by = "Inspection ID")
```

# CART Model

In this model, we first limit our focus to two types of inspections.  Either, the inspection was due to a regular Canvass, or the inspection was due to a complaint (which includes a possible food poisoning).  This eliminates other types of inspections, including re-inspections and licensing inspections. Next, we only consider restaurants which have had at least three regular Canvass inspections.

The first model is to predict which violations are major indicators of the result of an inspection. We consider the cases where the result can be either "Pass", "Pass with Conditions", "Fail".  We expect to see some violations are major red flags and will likely to cause a restaurants to Fail.


The second model runs on restaurants themselves.  We categorize restaurants into three groups:
- "Passes", where the restaurant has always received a "Pass" result;
- "Conditional", who have received at least one "Pass with Conditions"; and
- "Fails", where the restaurant has received at least one "Fail" result.

We restrict the inspections to those that are regular "Canvass" and not For Cause (That is, not do to complaint or suspected food poisoning), and require there be at least three "Canvass" inspections.


```{r Cart Libraries}
library(caret)
library(rpart)
library(rpart.plot)

#install.packages("rattle")
library(rattle)
```

## Complaint Classification
Here, we identify which restaurants have had a For Cause inspection (that is, due to a complaint or suspected food poisoning) at some point in there 

```{r}
canvass.inspection <- c("Canvass", "CANVASS", "CANVAS")
complaint.inspection <- c("Complaint", "Short Form Complaint", "Suspected Food Poisoning")


with.complaint <- inspections %>%
     dplyr::filter(`Inspection Type` %in% complaint.inspection)%>%
     dplyr::select(`License #`) %>%
     unique() %>%
     pull()

no.complaints <- inspections %>%
     dplyr::filter( !(`License #` %in% with.complaint)) %>%
     dplyr::filter(`Inspection Type` %in% canvass.inspection) %>%
     dplyr::select(`License #`) %>%
     unique() %>%
     pull()

inspections <- inspections %>%
     filter(
          (`License #` %in% with.complaint) | (`License #` %in% no.complaints)
     ) %>%
     mutate(Has_for_Cause = case_when(
          `License #` %in% with.complaint         ~ TRUE,
          `License #` %in% no.complaints          ~ FALSE
     ))

enough.canvass <- inspections %>%
     filter(`Inspection Type` %in% canvass.inspection) %>%
     group_by(`License #`) %>%
     summarise(NumberOfCanvass = n()) %>%
     filter(NumberOfCanvass > 2) %>%
     dplyr::select(`License #`) %>%
     distinct() %>%
     pull()


pass.fail.inspections <- inspections %>%
     filter(`License #` %in% enough.canvass,
            `License #` != 0,
            `Inspection Type` %in% c(canvass.inspection, complaint.inspection)) %>%
     filter(Results %in% c("Pass", "Fail", "Pass w/ Conditions")) %>%
     droplevels()

pass.fail.inspections <- pass.fail.inspections %>%
     mutate(Results = factor(Results, levels = c("Pass", "Pass w/ Conditions", "Fail")))

```

## Violations Per Inspection
This creates a column for each Violation Type. For each inspection, there is a 1 for each violation reported.

```{r}
pass.fail.inspections <- pass.fail.inspections %>%
     dplyr::select(`Inspection ID`, `License #`, `AKA Name`, `Facility Type`, 
                   `Inspection Date`, `Inspection Type`, `Results`, `Violations`) %>%
     mutate_at(vars(`Results`), factor)

violation_count <- inspect_long %>% group_by(`Inspection ID`, V_Num) %>%
     summarise(n = n()) %>%
     mutate(V_Num = str_pad(V_Num, 2, pad = "0")) %>%
     pivot_wider(names_from = V_Num,
                 names_prefix = "Violation_",
                 values_from = n,
                 values_fill = 0)

# Reorder the Columns in Violation Number Order
violation_count <- violation_count %>%
     dplyr::select(order(colnames(violation_count)))

pass.fail.inspections.wide <- pass.fail.inspections %>%
     left_join(violation_count, by = "Inspection ID")
```

## Pass/Fail Label and Violation Frequency
Here, we label restaurants as having failed at least once, or passed conditionally at least once, or passes always.

```{r}

# Get list of restaurants who have failed
restaurants.have.failed <- pass.fail.inspections.wide %>%
     filter(Results == "Fail") %>%
     dplyr::select(`License #`) %>%
     distinct(`License #`) %>%
     pull()

# Get list of restaurants who have passed conditionally

restaurants.passed.conditionally <- pass.fail.inspections.wide %>%
     filter(Results == "Pass w/ Conditions", 
            !(`License #` %in% restaurants.have.failed)) %>%
     dplyr::select(`License #`) %>%
     distinct(`License #`) %>%
     pull()

# Get list of restaurants who have alway passed

restaurants.always.pass <- pass.fail.inspections.wide %>%
     filter(Results == "Pass",
            !(`License #` %in% restaurants.have.failed),
            !(`License #` %in% restaurants.passed.conditionally)
            ) %>%
     dplyr::select(`License #`) %>%
     distinct(`License #`) %>%
     pull()

# Tag 

pass.fail.inspections.wide <- pass.fail.inspections.wide %>%
     mutate(Always.Results = case_when(
          `License #` %in% restaurants.have.failed          ~ "Fails",
          `License #` %in% restaurants.passed.conditionally ~ "Conditional",
          `License #` %in% restaurants.always.pass          ~ "Passes")
          )

ViolationNumbers <-  c(33, 34, 0, 32, 45, 39, 41, 35, 36, 38, 43, 14, 16, 30, 31, 
                       40, 11, 37, 3, 21, 19, 27, 42, 2, 18, 22, 29, 8, 24, 28, 
                       6, 44, 13, 25, 1, 9, 10, 12, 26, 70, 7, 17, 23, 4, 5, 15, 20)

ViolationNumbers <- str_pad(ViolationNumbers, 2, pad = "0")
ViolationColumns <- paste("Violation_", sort(ViolationNumbers), sep="")

restaurant.summary <- pass.fail.inspections.wide %>%
     group_by(`License #`) %>%
     summarize_at(ViolationColumns, mean) %>%
     mutate(Always.Results = case_when(
          `License #` %in% restaurants.have.failed          ~ "Fails",
          `License #` %in% restaurants.passed.conditionally ~ "Conditional",
          `License #` %in% restaurants.always.pass          ~ "Passes")
          ) %>%
     mutate(Always.Results = factor(
               Always.Results, 
               levels = c("Passes", "Conditional", "Fails"))
          )
```


## CART 1 - By Results
### Create Training Sets

```{r}
set.seed(482022)
trainIndex <- createDataPartition(pass.fail.inspections.wide$`Inspection ID`, p = .8, 
                                  list = FALSE, 
                                  times = 1)

inspections.train <- pass.fail.inspections.wide[trainIndex,]
inspections.test <- pass.fail.inspections.wide[-trainIndex,]
```

### Run CART
```{r Violations CART Model}
xFormula <- paste(ViolationColumns, collapse=' + ')
Formula <- as.formula(paste("Results ~ ", xFormula))

tic("Start CART Model")
model.cart1 <- rpart(Formula, data = inspections.train,
               control = c(
                    maxdepth = 4
               ))
toc()
```

### CART Analysis

```{r, fig.width = 8}
png("graphics/model_cart1_chart.png", width=7, height=6, units="in", res=144)
model_cart1_chart<-rpart.plot(model.cart1,
                  #box.palette = list('green','orange','red'))
                  box.palette = as.list(c(plasma(n=2, begin =0.45, end=0.8, direction=-1),plasma(n=1, begin=0.6, end=0.6))))
dev.off()

model_cart1_chart<-rpart.plot(model.cart1,
                  #box.palette = list('green','orange','red'))
                  box.palette = as.list(c(plasma(n=2, begin =0.45, end=0.8, direction=-1),plasma(n=1, begin=0.6, end=0.6))))

model.cart1$variable.importance
```

```{r}
cart1.test <-  predict(model.cart1, inspections.test, type = "class")

cart1.confusionMatrix <- confusionMatrix(data = cart1.test, reference = inspections.test$Results)
cart1.confusionMatrix
precision <- cart1.confusionMatrix$byClass['Pos Pred Value']    
recall <- cart1.confusionMatrix$byClass['Sensitivity']

```

## CART 2 - Long Term Violations 

### Create Train/Test Split

```{r}
set.seed(482022)
train2Index <- createDataPartition(restaurant.summary$`License #`, 
                                  p = .8, 
                                  list = FALSE, 
                                  times = 1)

restaurant.summary.train <- restaurant.summary[train2Index,]
restaurant.summary.test <- restaurant.summary[-train2Index,]
```
### Run CART

```{r}

Formula2 <- as.formula(paste("Always.Results ~ ", xFormula))

tic("Start CART Model")
model.cart2 <- rpart(Formula2, 
                data = restaurant.summary.train,
                control = c(
                    maxdepth = 4
               ))
toc()

```

### CART Analysis

```{r, fig.width=8}
png("graphics/model_cart2_chart.png", width=7, height=7, units="in", res=144)
rpart.plot(model.cart2,
           #box.palette = list('green','orange','red')
           box.palette = as.list(c(plasma(n=2, begin =0.45, end=0.8, direction=-1),plasma(n=1, begin=0.6, end=0.6)))
           )
dev.off()

rpart.plot(model.cart2,
           #box.palette = list('green','orange','red')
           box.palette = as.list(c(plasma(n=2, begin =0.45, end=0.8, direction=-1),plasma(n=1, begin=0.6, end=0.6)))
           )

model.cart2$variable.importance

```

```{r}
# cart2.test <-  predict(model.cart2, restaurant.summary.test, type = "class")
# table(cart2.test, restaurant.summary.test$Always.Results)

cart2.test <-  predict(model.cart2, restaurant.summary.test, type = "class")

cart2.confusionMatrix <- confusionMatrix(data = cart2.test, 
                                         reference = restaurant.summary.test$Always.Results)
cart2.confusionMatrix

precision <- cart2.confusionMatrix$byClass['Pos Pred Value']    
recall <- cart2.confusionMatrix$byClass['Sensitivity']
```





