---
title: "Demographic Maps"
output:
  html_document:
    df_print: paged
---

# Set Up

## Libraries

```{r echo=FALSE}
# Basic packages
library(plyr)
library(tidyverse)
library(ggplot2)
library(lubridate)

# Packages for mapping
library(sp)
library(raster)
library(sf)
library(leaflet)

# Formatting and Visualizations
library(viridis)          # Need for viridis color palette in word cloud
library(knitr)            # Need to manage text outputs in RMarkdown
library(formatR)          # To manage margins

```

## Importing Data

```{r, Importing-the-Data}
data_cols <- cols(
     `Inspection ID` = col_double(),
     `DBA Name` = col_character(),
     `AKA Name` = col_character(),
     `License #` = col_double(),
     `Facility Type` = col_character(),
     Risk = col_factor(),
     Address = col_character(),
     City = col_character(),
     State = col_character(),
     Zip = col_double(),
     `Inspection Date` = col_datetime(format = ""),
     `Inspection Type` = col_character(),
     Results = col_character(),
     Violations = col_character(),
     Latitude = col_double(),
     Longitude = col_double(),
     Location = col_character()
)
raw.data <- read_csv("../data/food-inspections.csv", col_types = data_cols)
```

## Communities

Chicago has community Areas and Neighborhoods. community areas have demographic data associated with them, but do not necessarily form homogeneous regions.

Shape files from [here](https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Neighborhoods/bbvz-uum9).

```{r Importing-Maps}
nbhd <-
  read_sf("../data/ChicagoNbhd/geo_export_8956960c-bd3c-4e92-a8b2-4748382c28a3.shp")  %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

communities <-
  read_sf("../data/ChicagoCommunities/geo_export_55b54106-a17e-4776-86c9-12db49043238.shp")  %>%
  sf::st_transform('+proj=longlat +datum=WGS84')
```

This creates a new df which has a GIS geometry column in place of Latitude and Longitude.

```{r Mapping-Inspections}
inspections <- raw.data  %>%
     filter(! (is.na(Latitude) | is.na(Longitude)))  %>%
     add_column(community = NA_character_, .after = "Address")  %>%
     rowid_to_column  %>%
     st_as_sf(
          coords = c("Longitude", "Latitude"), 
          crs= crs(nbhd))
```

## Demographics

The following is a table of demographic area based on community Area
For working with demographic fields, added a second table that has demographic information preceded by "POP_".  Master demographic table looks for a prefix

```{r}
demographic_names <- c(
     'Layer', 'Name', 'GEOID', 'Population', 'Longitude', 'Latitude', 'Male',
     'Female', 'White', 'Black', 'Asian', 'Hispanic', 'Native', 'TwoOrMore',
     'Infant', 'Senior', 'Adult', 'MiddleAge', 'Young', 'Children','Juvenile'
)

demographics_columns <- cols(
     Layer = col_character(),
     Name = col_character(),
     GEOID = col_character(),
     Population = col_double(),
     Longitude = col_character(),
     Latitude = col_character(),
     `Male` = col_double(),
     `Female` = col_double(),
     `White` = col_double(),
     `Black` = col_double(),
     `Asian` = col_double(),
     `Hispanic` = col_double(),
     `Native` = col_double(),
     `TwoOrMore` = col_double(),
     `Infant` = col_double(),
     `Senior` = col_double(),
     `Adult` = col_double(),
     `MiddleAge` = col_double(),
     `Young` = col_double(),
     `Children` = col_double(),
     `Juvenile` = col_double()
)

demographics.info <- read_csv("../data/Demographics/CHADDemographics.csv", 
                              col_names = demographic_names, 
                              skip = 1, 
                              n_max = 2,
                              show_col_types = FALSE) %>%
     dplyr::select('Male':'Juvenile')

demographics <- read_csv("../data/Demographics/CHADDemographics.csv",
                         col_names = demographic_names, 
                         col_types = demographics_columns,
                         skip = 5
)

# Renaming variables to ensure consistency across program paradigms.
demographic.columns.fixed <- str_c(rep("POPCT_", 15), 
                                   str_to_upper(demographic_names[7:21])) %>%
     append(demographic_names[1:6], after=0)

demographics <- demographics %>%
     setNames(demographic.columns.fixed) %>%
     mutate(Name = str_to_upper(Name))%>%
     as_tibble()

```

The table `demographicInfo` provides a description of the columns of `demographics`

Here, the `communities` and `demographics` dataframe are merged to create a `communities.demographics` dataframe.

```{r}
communities <- communities  %>%
     mutate(community = replace(community, community == "OHARE", "O'HARE"))

demographics <- demographics  %>%
     mutate(commID = str_to_upper(Name))

communities.demographics <- merge(communities, 
                                  demographics, 
                                  by.x = "community", 
                                  by.y = "commID")
```

```{r other_demographics}
demo.traf.vac <-
  read.csv("../data/Demographics/Chicago_traffic_and_Vacant.csv") %>% 
  mutate(Name = str_to_upper(Name))
demo.food.act <-
  read.csv("../data/Demographics/Chicago_food_and_activity.csv") %>% 
  mutate(Name = str_to_upper(Name))
demo.health <-
  read.csv("../data/Demographics/Chicago_Health_Status.csv") %>% 
  mutate(Name = str_to_upper(Name))
demo.income <-
  read.csv("../data/Demographics/Chicago_income.csv") %>% 
  mutate(Name = str_to_upper(Name))
demo.lang.birth <-
  read.csv("../data/Demographics/Chicago_lang_and_birth.csv") %>% 
  mutate(Name = str_to_upper(Name))
demo.morbidity <-
  read.csv("../data/Demographics/Chicago_Morbidity_Rate.csv") %>% 
  mutate(Name = str_to_upper(Name))
demo.mortality <-
  read.csv("../data/Demographics/Chicago_mortality.csv") %>% 
  mutate(Name = str_to_upper(Name))
demo.population <-
  read.csv("../data/Demographics/Chicago_pop.csv") %>% 
  mutate(Name = str_to_upper(Name))
```

```{r demo_graphs, fig.width=12, fig.height=12}
ggplot(communities.demographics) + geom_sf(aes(fill = as.integer(POPCT_WHITE)))
ggplot(communities.demographics) + geom_sf(aes(fill = as.integer(POPCT_BLACK)))
ggplot(communities.demographics) + geom_sf(aes(fill = as.integer(POPCT_ASIAN)))
ggplot(communities.demographics) + geom_sf(aes(fill = as.integer(POPCT_HISPANIC)))
ggplot(communities.demographics) + geom_sf(aes(fill = as.integer(POPCT_SENIOR)))
ggplot(communities.demographics) + geom_sf(aes(fill = as.integer(POPCT_ADULT)))
ggplot(communities.demographics) + geom_sf(aes(fill = as.integer(POPCT_CHILDREN)))

ggplot(inner_join(demo.traf.vac, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = TRF_2018, geometry = geometry))
ggplot(inner_join(demo.traf.vac, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = TRF_2017, geometry = geometry))
ggplot(inner_join(demo.traf.vac, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = TRF_2016, geometry = geometry))
ggplot(inner_join(demo.traf.vac, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = VAC_2014.2018, geometry = geometry))
ggplot(inner_join(demo.traf.vac, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = VAC_2013.2017, geometry = geometry))
ggplot(inner_join(demo.traf.vac, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = VAC_2012.2016, geometry = geometry))
ggplot(inner_join(demo.traf.vac, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = VAC_2011.2015, geometry = geometry))
ggplot(inner_join(demo.traf.vac, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = VAC_2010.2014, geometry = geometry))
ggplot(inner_join(demo.income, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = PCI_2014.2018, geometry = geometry))
ggplot(inner_join(demo.income, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = PCI_2013.2017, geometry = geometry))
ggplot(inner_join(demo.income, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = PCI_2012.2016, geometry = geometry))
ggplot(inner_join(demo.income, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = PCI_2011.2015, geometry = geometry))
ggplot(inner_join(demo.income, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = PCI_2010.2014, geometry = geometry))
ggplot(inner_join(demo.income, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = POV_2014.2018, geometry = geometry))
ggplot(inner_join(demo.income, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = POV_2013.2017, geometry = geometry))
ggplot(inner_join(demo.income, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = POV_2012.2016, geometry = geometry))
ggplot(inner_join(demo.income, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = POV_2011.2015, geometry = geometry))
ggplot(inner_join(demo.health, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = HCSOHS_2016.2018, geometry = geometry))
ggplot(inner_join(demo.health, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = HCSOHS_2015.2017, geometry = geometry))
ggplot(inner_join(demo.health, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = HCSOHS_2014.2016, geometry = geometry))
ggplot(inner_join(demo.health, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = HCSOHSP_2016.2018, geometry = geometry))
ggplot(inner_join(demo.health, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = HCSOHSP_2015.2017, geometry = geometry))
ggplot(inner_join(demo.health, communities, by = c("Name" = "community"))) +
  geom_sf(aes(fill = HCSOHSP_2014.2016, geometry = geometry))
```

# Cleaning Data

## Labeling Violations with Community

We look for the intersections of `communities` and `inspection` and create `vio.comm.int`, which is a sparse matrix. For a Community Area descriped in row `i` of `communities` there is a row in `vio.comm.int` which lists the rows `j` in `inspections` which fall within its boundary.

```{r}
vio.comm.int <- st_intersects(communities, inspections)
```

Now, we revise `inspections` to include a column listing the Community Area the location falls within.

```{r message=FALSE, warning=FALSE}
inspections <- inspections  %>% as_tibble()
for (i in 1:nrow(communities)) {
     inspections <- inspections %>%
          rows_update(
               tibble(
                    rowid = unlist(vio.comm.int[[i]]), 
                    community = communities$community[i]
               )
          )
}
inspections <- inspections  %>% st_as_sf()
```

## Risk

```{r}
inspections <- inspections  %>%
     mutate(
          Risk = fct_recode(inspections$Risk,
               Low = "Risk 3 (Low)", 
               Medium = "Risk 2 (Medium)", 
               High = "Risk 1 (High)")) 
```

## Facility Type

```{r}
#For consistency, first convert facilities to lowercase.
#Then, limit to inspection dates before paradigm changed.
#Look for any odd cases where a facility contains restaurant in name or is misspelled.  
inspections <- inspections  %>%
     mutate(`Facility Type` = str_to_lower(`Facility Type`)) %>%   
     dplyr::filter(`Inspection Date`<"2018-07-01", 
                   `Facility Type`=="restaurant")
```
# Exploration Tools

## Remove Geometry

Geometry is dropped from here on out to simplify issues.

```{r data_reshape}
#keep copy for calculation purposes
inspections.geo<-inspections

#drop geometry for simpicity of inspection.  
inspections.noGeo <- st_drop_geometry(inspections)
```

```{r biz_summary, fig.width=12, fig.height = 12}
#How many businesses are in each community
biz_by_community<-inspections.noGeo%>%dplyr::select(community, `License #`, `AKA Name`)%>%distinct()%>%group_by(community)%>%dplyr::summarize(biz.ct=n())

#Business Information Table
biz.info<-inspections%>%
  dplyr::select(-c(City, State, Zip, `Inspection Type`, `Inspection Date`, Results, Violations))%>%
  distinct()

# How many years were each business inspected?
# Use this in downstream rate calculations
# #broken for license #0, which impacts multiple businesses.
# Include AKA Name to avoid conflicts where multiple businesses fall under same license
yrs_by_biz<-inspections.noGeo%>%
  dplyr::select(`License #`,`AKA Name`,community,`Inspection Date`)%>%
  mutate(`Inspection Date`=year(`Inspection Date`))%>%
  distinct()%>%group_by(`License #`, `AKA Name`, community)%>%
  dplyr::summarize(yr.ct=n())

#How mnany times was a business inspected?
inspect_by_biz<-inspections.noGeo%>%
  dplyr::select(`License #`, `AKA Name`,community)%>%
  group_by(`License #`, `AKA Name`, community)%>%
  dplyr::summarize(biz.insp.ct=n())

#Calculate the inspections per business per year
inspect_per_biz_per_yr<-inner_join(yrs_by_biz, inspect_by_biz)%>%
  mutate(biz.insp.rate=biz.insp.ct/yr.ct)

#What is the average inspection rate in a community?
inspect_rate_by_cmty<-inspect_per_biz_per_yr%>%
  dplyr::select(community, biz.insp.rate)%>%
  group_by(community)%>%
  dplyr::summarise(yr.cmty.insp.rate=mean(biz.insp.rate))

# new tibble with inspection rate and business info combined
inspect_per_biz_per_yr_w_geom<-inner_join(biz.info, inspect_per_biz_per_yr)

```

```{r biz_graphs}
#Show the number of businesses in a community
inner_join(communities, biz_by_community, by = c("community" = "community")) %>%
  ggplot() +
  geom_sf(aes(fill = biz.ct)) +
  scale_fill_viridis_c(option = "C", name = "Count") + 
  labs(title = "Businesses per Community")

#Show the community inspection rate
inner_join(communities,
           inspect_rate_by_cmty,
           by = c("community" = "community")) %>%
  ggplot() +
  geom_sf(aes(fill = yr.cmty.insp.rate)) +
  scale_fill_viridis_c(option = "C", name = "Inspections") + 
  labs(title = "Average Inspections per Business per Year by Community")

#Show an overlay of business inspection frequency
inner_join(communities,
           inspect_rate_by_cmty,
           by = c("community" = "community")) %>%
  ggplot() + geom_sf(aes(fill = yr.cmty.insp.rate)) +
  geom_sf(data = inspect_per_biz_per_yr_w_geom, aes(color = biz.insp.rate)) +
  scale_fill_viridis_c(option = "C", name = "Cmty. Inspecitons") +
  scale_color_viridis_c(option = "C", name = "Bus. Inspections") + 
  labs(title = "Business Inspection Frequency vs. Community Inspection Frequency)")

#summarize mean years in business for community.
cmty_biz_yrs <- inspect_per_biz_per_yr %>%
  group_by(community) %>%
  dplyr::select(community, yr.ct) %>%
  dplyr::summarise(cmty.biz.yrs = mean(yr.ct))

#show mean years in business for community.
inner_join(communities, cmty_biz_yrs, by = c("community" = "community")) %>%
  ggplot() +
  geom_sf(aes(fill = cmty.biz.yrs)) +
  scale_fill_viridis_c(option = "C", name = "Age (Years)") +
  labs(title = "Average Business Age by Community")
  
```

```{r output_key_graphs}

png("../img/biz_ct_map.png", width = 6.5, height = 6.5, units ="in", res=144)

inner_join(communities, biz_by_community, by = c("community" = "community")) %>%
  ggplot() +
  geom_sf(aes(fill = biz.ct)) +
  scale_fill_viridis_c(option = "C", name = "Count") + 
  labs(title = "Businesses per Community")

dev.off()

png("../img/demo_pop.png", width = 6.5, height = 6.5, units ="in", res=144)
ggplot(communities.demographics, ) + 
  geom_sf(aes(fill = as.integer(Population))) +
    scale_fill_viridis_c(option = "C", name = "Population") + 
  labs(title = "Population by Community") 

dev.off()

png("../img/median_income.png", width = 6.5, height = 6.5, units ="in", res=144)
inner_join(communities, demo.income, by = c("community" = "Name")) %>%
  ggplot() +
  geom_sf(aes(fill = PCI_2014.2018)) +
  scale_fill_viridis_c(option = "C", name = "Dollars") + 
  labs(title = "Per Capita Income, 2014 - 2018") 

dev.off()

# Probanility two residents chosen at random will belong to different family income levels
png("../img/economic_diversity.png", width = 6.5, height = 6.5, units ="in", res=144)
inner_join(communities, demo.income, by = c("community" = "Name")) %>%
  ggplot() +
  geom_sf(aes(fill = EDX_2014.2018)) +
  scale_fill_viridis_c(option = "C", name = "Probability") + 
  labs(title = "Economic Diversity Index, 2014 - 2018") 
dev.off()

```