---
title: "CART Model"
output:
  html_document:
    df_print: paged
---


# Set Up

## Libraries

```{r echo=FALSE}
# Basic packages
#library(plyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
library(tictoc)

# Packages for mapping
library(sp)
library(sf)
library(raster)
library(leaflet)

# Formatting and Visualizations
library(viridis)          # Need for viridis color palette in word cloud
library(knitr)            # Need to manage text outputs in RMarkdown
library(formatR)          # To manage margins

```

## Importing Data

```{r, Importing-the-Data}
data_cols <- cols(
     `Inspection ID` = col_double(),
     `DBA Name` = col_character(),
     `AKA Name` = col_character(),
     `License #` = col_double(),
     `Facility Type` = col_character(),
     Risk = col_factor(),
     Address = col_character(),
     City = col_character(),
     State = col_character(),
     Zip = col_double(),
     `Inspection Date` = col_datetime(format = ""),
     `Inspection Type` = col_character(),
     Results = col_character(),
     Violations = col_character(),
     Latitude = col_double(),
     Longitude = col_double(),
     Location = col_character()
)
raw.data <- read_csv("../data/food-inspections.csv", col_types = data_cols)
```

## Communities

Chicago has community Areas and Neighborhoods. community areas have demographic data associated with them, but do not necessarily form homogeneous regions.

Shape files from [here](https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Neighborhoods/bbvz-uum9).

```{r Importing-Maps}
nbhd <- read_sf("../data/ChicagoNbhd/geo_export_8956960c-bd3c-4e92-a8b2-4748382c28a3.shp")  %>%
     sf::st_transform('+proj=longlat +datum=WGS84')

communities <- read_sf("../data/ChicagoCommunities/geo_export_55b54106-a17e-4776-86c9-12db49043238.shp")  %>%
     sf::st_transform('+proj=longlat +datum=WGS84')
```

This creates a new df which has a GIS geometry column in place of Latitude and Longitude.

```{r Mapping-Inspections}
inspections <- raw.data  %>%
     filter(! (is.na(Latitude) | is.na(Longitude)))  %>%
     add_column(community = NA_character_, .after = "Address")  %>%
     rowid_to_column  %>%
     st_as_sf(
          coords = c("Longitude", "Latitude"), 
          crs= crs(nbhd))
```

## Demographics

The following is a table of demographic area based on community Area
For working with demographic fields, added a second table that has demographic information preceded by "POP_".  Master demographic table looks for a prefix

```{r}
demographic_names <- c(
     'Layer', 'Name', 'GEOID', 'Population', 'Longitude', 'Latitude', 'Male',
     'Female', 'White', 'Black', 'Asian', 'Hispanic', 'Native', 'TwoOrMore',
     'Infant', 'Senior', 'Adult', 'MiddleAge', 'Young', 'Children','Juvenile'
)

demographics_columns <- cols(
     Layer = col_character(),
     Name = col_character(),
     GEOID = col_character(),
     Population = col_double(),
     Longitude = col_character(),
     Latitude = col_character(),
     `Male` = col_double(),
     `Female` = col_double(),
     `White` = col_double(),
     `Black` = col_double(),
     `Asian` = col_double(),
     `Hispanic` = col_double(),
     `Native` = col_double(),
     `TwoOrMore` = col_double(),
     `Infant` = col_double(),
     `Senior` = col_double(),
     `Adult` = col_double(),
     `MiddleAge` = col_double(),
     `Young` = col_double(),
     `Children` = col_double(),
     `Juvenile` = col_double()
)

demographics.info <- read_csv("../data/Demographics/CHADDemographics.csv", 
                              col_names = demographic_names, 
                              skip = 1, 
                              n_max = 2,
                              show_col_types = FALSE) %>%
     dplyr::select('Male':'Juvenile')

demographics <- read_csv("../data/Demographics/CHADDemographics.csv",
                         col_names = demographic_names, 
                         col_types = demographics_columns,
                         skip = 5
)

# Renaming variables to ensure consistency across program paradigms.
demographic.columns.fixed <- str_c(rep("POPCT_", 15), str_to_upper(demographic_names[7:21])) %>%
     append(demographic_names[1:6], after=0)

demographics <- demographics %>%
     setNames(demographic.columns.fixed) %>%
     mutate(Name = str_to_upper(Name))%>%
     as_tibble()

```

The table `demographicInfo` provides a description of the columns of `demographics`

Here, the `communities` and `demographics` dataframe are merged to create a `communities.demographics` dataframe.

```{r}
communities <- communities  %>%
     mutate(community = replace(community, community == "OHARE", "O'HARE"))

demographics <- demographics  %>%
     mutate(commID = str_to_upper(Name))

communities.demographics <- merge(communities, 
                                  demographics, 
                                  by.x = "community", 
                                  by.y = "commID")
```


```{r other_demographics, eval=TRUE, include=TRUE}
demo.traf.vac <- read.csv("../data/Demographics/Chicago_traffic_and_Vacant.csv") %>%
     mutate(Name=str_to_upper(Name))

demo.income <- read.csv("../data/Demographics/Chicago_income.csv") %>% 
     mutate(Name=str_to_upper(Name))

demo.lang.birth <- read.csv("../data/Demographics/Chicago_lang_and_birth.csv") %>%
     mutate(Name=str_to_upper(Name))

demo.population <- read.csv("../data/Demographics/Chicago_pop.csv") %>% 
     mutate(Name=str_to_upper(Name))
```

# Cleaning Data

## Labeling Violations with community

We look for the intersections of `communities` and `inspection` and create `vio.comm.int`, which is a sparse matrix. For a community Area described in row `i` of `communities` there is a row in `vio.comm.int` which lists the rows `j` in `inspections` which fall within its boundary.

```{r Intersecting-Inspections-Communities}
tic("Running st_intersects")
vio.comm.int <- st_intersects(communities, inspections)
toc()
```

Now, we revise `inspections` to include a column listing the community Area the location falls within.

```{r Label Inspection by Community Area}
tic("Label Inspections by Community Area")
inspections <- inspections  %>% as_tibble()
for (i in 1:nrow(communities)) {
     inspections <- inspections %>%
          rows_update(
               tibble(
                    rowid = unlist(vio.comm.int[[i]]), 
                    community = communities$community[i]
               ),
               by = "rowid"
          )
}
inspections <- inspections  %>% st_as_sf()
toc()
```

## Risk
Update the risk category to Low / Medium / High
```{r}
inspections <- inspections  %>%
     mutate(
          Risk = fct_recode(inspections$Risk,
                            Low = "Risk 3 (Low)", 
                            Medium = "Risk 2 (Medium)", 
                            High = "Risk 1 (High)")) 
```

## Facility Type

```{r}
#For consistency, first convert facilities to lowercase.
#Then, limit to inspection dates before paradigm changed.
#Look for any odd cases where a facility contains restaurant in name or is misspelled.  
inspections <- inspections  %>%
     mutate(`Facility Type` = str_to_lower(`Facility Type`)) %>%   
     dplyr::filter(`Inspection Date`<"2018-07-01", 
                   `Facility Type`=="restaurant")
```

## Violation Field

Collect subset of violations.  

```{r}
#pre: Inspection data is filtered to prior to 2018-07-01
#pre: Inspection is filtered to facility == restaurant.  
tic("Splitting Violations")

inspections <- inspections  %>% 
     mutate(
          `Violations` = if_else(is.na(`Violations`), 
                                 true =  "0. NO VIOLATION RECORDED - Comments: NO VIOLATION RECORDED", 
                                 false = `Violations`),
     )

violations <- inspections  %>%
     dplyr::select(c("Inspection ID", "Violations"))  %>%
     separate_rows(Violations, sep="\\|")  %>%
     separate(Violations, c("Violation", "Observation"), 
              sep = "- Comments: ",
              fill = "right") %>% # Fill makes empty comments without complaining
     separate(Violation, c("V_Num", "V_Desc"), 
              sep = "\\.\\s",
              extra="merge" # Only split on the first . 
     )
violations <- violations %>%
     mutate(
          V_Num = str_replace_all(violations$V_Num,"\\s",""),
          V_Desc = str_replace_all(V_Desc,"^\\s",""),
          V_Desc = str_replace_all(V_Desc,"\\s$",""),
          V_Num = as.integer(V_Num))
toc()
```

## Remove Geometry

Geometry is dropped from here on out to simplify issues.

```{r data_reshape}
#keep copy for calculation purposes
inspections.geo<-inspections
#drop geometry for simpicity of inspection.  
violations <- st_drop_geometry(violations)
inspections <- st_drop_geometry(inspections)

#inspect_long is key combination for downstream calcs and figs
inspect_long <- inspections %>%
     dplyr::select(-c(Violations, Location))%>%
     inner_join(violations, by = "Inspection ID")
```

```{r}
canvass.inspection <- c("Canvass", "CANVASS", "CANVAS")
complaint.inspection <- c("Complaint", "Short Form Complaint", "Suspected Food Poisoning")


with.complaint <- inspections %>%
     dplyr::filter(`Inspection Type` %in% complaint.inspection)%>%
     dplyr::select(`License #`) %>%
     unique() %>%
     pull()

no.complaints <- inspections %>%
     dplyr::filter( !(`License #` %in% with.complaint)) %>%
     dplyr::filter(`Inspection Type` %in% canvass.inspection) %>%
     dplyr::select(`License #`) %>%
     unique() %>%
     pull()

inspections <- inspections %>%
     filter(
          (`License #` %in% with.complaint) | (`License #` %in% no.complaints)
     ) %>%
     mutate(Has_for_Cause = case_when(
          `License #` %in% with.complaint ~ TRUE,
          `License #` %in% no.complaints ~ FALSE
     ))

enough.canvass <- inspections %>%
     filter(`Inspection Type` %in% canvass.inspection) %>%
     group_by(`License #`) %>%
     summarise(NumberOfCanvass = n()) %>%
     filter(NumberOfCanvass > 2) %>%
     dplyr::select(`License #`) %>%
     distinct() %>%
     pull()


pass.fail.inspections <- inspections %>%
     filter(`License #` %in% enough.canvass,
            `License #` != 0,
            `Inspection Type` %in% c(canvass.inspection, complaint.inspection)) %>%
     filter(Results %in% c("Pass", "Fail", "Pass w/ Conditions")) %>%
     droplevels()

pass.fail.inspections <- pass.fail.inspections %>%
     mutate(Results = factor(Results, levels = c("Pass", "Pass w/ Conditions", "Fail")))

```
## Violations Per Inspection
This creates a column for each Violation Type. For each inspection, there is a 1 for each violation reported.

```{r}
pass.fail.inspections <- pass.fail.inspections %>%
     dplyr::select(`Inspection ID`, `License #`, `AKA Name`, `Facility Type`, 
                   `Inspection Date`, `Inspection Type`, `Results`, `Violations`) %>%
     mutate_at(vars(`Results`), factor)

violation_count <- inspect_long %>% group_by(`Inspection ID`, V_Num) %>%
     summarise(n = n()) %>%
     mutate(V_Num = str_pad(V_Num, 2, pad = "0")) %>%
     pivot_wider(names_from = V_Num,
                 names_prefix = "Violation_",
                 values_from = n,
                 values_fill = 0)

# Reorder the Columns in Violation Number Order
violation_count <- violation_count %>%
     dplyr::select(order(colnames(violation_count)))

pass.fail.inspections.wide <- pass.fail.inspections %>%
     left_join(violation_count, by = "Inspection ID")
```

#Stats
```{r}
total_inspections<-inspections%>%
     mutate(year=year(`Inspection Date`))%>%
     filter(year<2018)%>%
     group_by(year)%>%
     dplyr::summarize(insp.yr=n())%>%
     dplyr::summarize(insp.yr=mean(insp.yr))

yearly_inspections<-inspections%>%
     mutate(year=year(`Inspection Date`))%>%
     filter(year<2018)%>%
     group_by(year)%>%
     dplyr::select(year, Results)%>%
     group_by(year, Results)%>%
     dplyr::summarise(insp.res.ct=n())%>%
     group_by(Results)%>%
     dplyr::summarize(mean_types=mean(insp.res.ct))

cmty_inspections<-inspections%>%
     dplyr::select(community,`Inspection ID`)%>%
     group_by(community)%>%
     dplyr::summarise(insp.ct=n())

total_violations<-inspect_long%>%
     mutate(year=year(`Inspection Date`))%>%
     filter(year<2018)%>%
     group_by(year)%>%
     dplyr::select(year, V_Num)%>%
     group_by(year)%>%
     dplyr::summarise(v.ct=n())%>%
     dplyr::summarize(mean_types=mean(v.ct))

violations_by_type<-inspect_long%>%
     mutate(year=year(`Inspection Date`))%>%
     filter(year<2018)%>%
     dplyr::select(V_Num)%>%
     group_by(V_Num)%>%
     dplyr::summarise(v.ct=n())%>%
     mutate(v_pct=round(v.ct/sum(v.ct)*100, 2))

#Summaries violations per inspection per inspection type
viol_by_insp_type<-inspect_long%>%
     group_by(`Inspection Type`)%>%
     dplyr::summarise(tot_viol=n())

viol_per_inspect_aka <-
     inspect_long %>%
     dplyr::select(`Inspection ID`, `AKA Name`,`License #`,
                   `Facility Type`, Risk, `Inspection Date`, V_Num) %>%
     group_by(`License #`,`AKA Name`, `Inspection ID`,`Facility Type`,
              `Inspection Date`, Risk) %>%
     dplyr::summarise(n = n())

viol_mean_per_aka <- viol_per_inspect_aka %>%
     group_by(`License #`, `AKA Name`, `Facility Type`, Risk) %>%
     dplyr::summarise(mean = mean(n), n_inspect = n())

total_restaurants<-inspections%>%
     mutate(year=year(`Inspection Date`))%>%
     filter(year<2018)%>%
     dplyr::select(`License #`, `AKA Name`)%>%
     distinct()%>%
     dplyr::summarize(rest_ct=n())


```
#Violation Description

```{r Violation_Description, fig.width=18}

#Key violation numbers from CART model
V_int<-c(3, 18, 21, 24, 28)

vio_pct_as_key.plot <- violations%>%
     dplyr::select(V_Num)%>%
     group_by(V_Num)%>%
     summarize(n=n())%>%
     mutate(n_pct=n/sum(n)*100) %>%
     mutate(keyViolation = as.logical(V_Num %in% V_int)) %>%
     ggplot()+
     geom_col(aes(y=as.factor(V_Num), 
                  x=n_pct, 
                  fill=keyViolation))+
     #scale_fill_manual(values = palette.colors(palette="ggplot2")[c(2,4)])+
     scale_fill_manual(values=plasma(n=2, begin =0.2, end=0.8))+
     labs(title="Percentage of Violations Observed in Businesses",
          subtitle="Usage in Classification Model Highlighted", 
          fill="Key Violation", 
          y="violation number", 
          x="Percent Observed")
vio_pct_as_key.plot

png("../img/viol_pct_as_key.png", width = 6.5, height = 8, units="in", res=144)
vio_pct_as_key.plot
dev.off()
```


```{r}
viol_pct_prsnt.plot <- violations%>%
     dplyr::select(V_Num)%>%
     group_by(V_Num)%>%
     summarize(n=n())%>%
     mutate(n_pct=n/sum(n)*100)%>%
     ggplot()+
     #geom_col(aes(x=as.factor(V_Num), y=n_pct), fill=palette.colors(palette="ggplot2")[c(4)])+
      geom_col(aes(x=as.factor(V_Num), y=n_pct), fill=plasma(n=1))+
     labs(title="Percentage of Violations Observed in Businesses",
          fill="Key Violation", 
          x="violation number", 
          y="Percent Observed")

viol_pct_prsnt.plot

png("../img/viol_pct_prsnt.png", width = 9, height = 6.5, units="in", res=144)
viol_pct_prsnt.plot
dev.off()
```



```{r}
#look at just the violations of interest
#what is the division with business category
key_viol_pct.plot <- violations%>%
     dplyr::select(V_Num, `Inspection ID`)%>%
     dplyr::filter(V_Num %in% V_int)%>%
     inner_join(inspect_long)%>%
     dplyr::filter(Results %in% c("Pass", "Pass w/ Conditions", "Fail"))%>%
     mutate(Results=factor(Results, c("Pass w/ Conditions", "Fail", "Pass")))%>%
     group_by(V_Num, Results)%>%
     summarize(n=n())%>%
     mutate(n_pct=n/sum(n)*100)%>%
     ggplot()+
     geom_col(aes(y=as.factor(V_Num), x=n_pct, fill=Results), 
              position = position_dodge(preserve="single"), 
              orientation="y")+
     scale_fill_manual(values=plasma(n=3, begin =0.2, end=0.8, direction = 1))+
     #scale_fill_manual(values=c("orange", "red", "green"))+
     labs(title="Violations Used in Classification Model",
          subtitle="as Percentage of Total Observations vs Inspection Outcome",
          fill="Worst Inspection Outcome",
          y="Violation Number", 
          x="Percent Observed")
key_viol_pct.plot

png("../img/key_viol_pct.png", width=6.5, height=6.5, units="in", res=144)
key_viol_pct.plot
dev.off()
```



```{r}
#look at just the violations of interest
#what is the division with business category
key_viol_pct_prsnt.plot <- violations%>%
     dplyr::select(V_Num, `Inspection ID`)%>%
     dplyr::filter(V_Num %in% V_int)%>%
     inner_join(inspect_long)%>%
     dplyr::filter(Results %in% c("Pass", "Pass w/ Conditions", "Fail"))%>%
     mutate(Results=factor(Results, c("Pass w/ Conditions", "Fail", "Pass")))%>%
     group_by(V_Num, Results)%>%
     summarize(n=n())%>%
     mutate(n_pct=n/sum(n)*100)%>%
     ggplot()+
     geom_col(aes(x=as.factor(V_Num), y=n_pct, fill=Results), 
              position = position_dodge(preserve="single"), 
              orientation="x")+
     scale_fill_manual(values=plasma(n=3, begin =0.2, end=0.8, direction = 1))+
     #scale_fill_manual(values=c("orange", "red", "green"))+
     labs(title="Violations Used in Classification Model",
          subtitle="as Percentage of Total Observations vs Inspection Outcome",
          fill="Worst Inspection Outcome",
          x="Violation Number", 
          y="Percent Observed")
key_viol_pct_prsnt.plot
png("../img/key_viol_pct_prsnt.png", width=9, height=6.5, units="in", res=144)
key_viol_pct_prsnt.plot
dev.off()
```


```{r, fig.height=12}
#look at all violations 
#For each violation
#what is the percentage of each pass/fail/conditional category receiving it? 

key_viol_pct_w_class.data <- violations%>%
     dplyr::select(V_Num, `Inspection ID`)%>%
     inner_join(inspect_long)%>%
     dplyr::filter(Results %in% c("Pass", "Pass w/ Conditions", "Fail"))%>%
     mutate(Results=factor(Results, c("Pass w/ Conditions", "Fail", "Pass")))%>%
     group_by(V_Num, Results)%>%
     summarize(n=n())%>%
     mutate(n_pct=n/sum(n)*100)

data1 <- key_viol_pct_w_class.data %>%
     filter(V_Num %in% V_int)

data2 <- key_viol_pct_w_class.data %>%
     filter(!(V_Num %in% V_int))

key_viol_pct_w_class.plot <- ggplot(data = key_viol_pct_w_class.data)+
     geom_col(data = data1,
              aes(y=as.factor(str_pad(V_Num,2, side="left")), x=n_pct, fill=as.factor(Results)),
              color = "black",
              size=1) +
     geom_col(data = data2,
              aes(y=as.factor(str_pad(V_Num,2, side="left")), x=n_pct, fill=as.factor(Results)),
              size=1) +     
     scale_color_discrete(type=c("white", "black"))+
     scale_fill_manual(values=plasma(n=3, begin =0.2, end=0.8, direction = 1))+
     #scale_fill_manual(values=c("orange", "red", "green"))+
     labs(title="Violations as Percentage of Total Observations",
          subtitle="by Inspection Outcome and Use in Classificaiton Model",
          fill="Worst Inspection Outcome",
          color="Key Variable",
          y="Violation Number", 
          x="Percent Observed")+
      theme(legend.position = ("bottom"))
key_viol_pct_w_class.plot

png("../img/key_viol_pct_w_class.png", width = 6.5, height = 7.5, units="in", res=144)
key_viol_pct_w_class.plot
dev.off()
```


```{r}
png("../img/key_viol_pct_w_class_prsnt.png", width=18, height=6, units="in", res = 144)
violations%>%
     dplyr::select(V_Num, `Inspection ID`)%>%
     inner_join(inspect_long)%>%
     dplyr::filter(Results %in% c("Pass", "Pass w/ Conditions", "Fail"))%>%
     group_by(V_Num, Results)%>%
     summarize(n=n())%>%
     mutate(n_pct=n/sum(n)*100)%>%
     ggplot()+
     geom_col(aes(x=as.factor(V_Num), y=n_pct, fill=as.factor(Results), color=as.factor(V_Num%in%V_int)),
              size=1, 
              position = position_dodge(preserve="single"),                                                                                                        orientation="x")+
     scale_color_discrete(type=c("grey", "black"))+
     labs(title="Violations as Percentage of Total Observations",
          subtitle="by Inspection Outcome and Use in Classificaiton Model",
          fill="Worst Inspection Outcome",
          color="Key Variable",
          x="violation number", 
          y="Percent Observed")
dev.off()
```
#Data Set Observations
```{r}
n2<-inspections%>%dplyr::select(`AKA Name`, `License #`, Address)%>%
     distinct()%>%
     group_by(`License #`)%>%dplyr::summarize(n=n())%>%filter(n>2)
```

## community Business Summary Tables
```{r summarized_maps, fig.width=12, fig.height = 12}
tot_biz<-inspections%>%
     dplyr::select(`License #`, `AKA Name`)%>%
     distinct()%>%
     #group_by(`License #`)%>%
     dplyr::summarise(n=n()) 

#How many businesses are in each community
biz_by_community<-inspections%>%
     dplyr::select(community, `License #`, `AKA Name`)%>%
     distinct()%>%
     group_by(community)%>%
     summarize(biz.ct=n())

#Business Information Table
biz.info<-inner_join(inspections, communities)%>%
     dplyr::select(-c(City, State, Zip, `Inspection Type`, `Inspection Date`, Results, Violations))%>%
     distinct()

# How many years were each business inspected?
# Use this in downstream rate calculations
# #broken for license #0, which impacts multiple businesses.
# Include AKA Name to avoid conflicts where multiple businesses fall under same license
yrs_by_biz<-inspections%>%
     dplyr::select(`License #`,`AKA Name`,community,`Inspection Date`)%>%
     mutate(`Inspection Date`=year(`Inspection Date`))%>%
     distinct()%>%group_by(`License #`, `AKA Name`, community)%>%
     summarize(yr.ct=n())

#summarize mean years in business for community.
cmty_biz_yrs<-yrs_by_biz%>%
     group_by(community)%>%
     dplyr::select(community, yr.ct)%>%
     summarise(cmty.biz.yrs=mean(yr.ct))
```

###Comunity Business Inspection Rate Summary
```{r com_biz_sum}
#How many times was a business inspected?
inspect_by_biz<-inspections%>%
     dplyr::select(`License #`, `AKA Name`,community)%>%
     group_by(`License #`, `AKA Name`, community)%>%
     summarize(biz.insp.ct=n())

#Calculate the inspections per business per year
inspect_per_biz_per_yr<-inner_join(yrs_by_biz, inspect_by_biz)%>%
     mutate(biz.insp.rate=biz.insp.ct/yr.ct)

#What is the average inspection rate in a community?
inspect_rate_by_cmty<-inspect_per_biz_per_yr%>%
     dplyr::select(community, biz.insp.rate)%>%
     group_by(community)%>%
     summarise(yr.cmty.insp.rate=mean(biz.insp.rate))

# new tibble with inspection rate and business info combined
inspect_per_biz_per_yr_w_geom<-inner_join(biz.info, inspect_per_biz_per_yr)

#summary stat.  community violation rate per inspection
#this is interesting.  Look at the higher number of violations in outer areas.
cmty_viol_per_inspect<-inner_join(inspections, filter(viol_mean_per_aka, n_inspect>2))%>%
     dplyr::select(community, mean)%>%
     group_by(community)%>%
     summarise(cmty_viol_rate=mean(mean))

cmty_results_per_inspect<-inner_join(inspections, filter(viol_mean_per_aka, n_inspect>2))%>%
     dplyr::select(community, mean)%>%
     group_by(community)%>%
     summarise(cmty_viol_rate=mean(mean))
```

###Graphs of community business summaries

```{r community_biz_plot}

##cmty_stat_plot() graphs layered community data with aesthetic fill:
##pre: community is defined sf shape
##pre: data is defined summary statistic containing a 'community' name 
## and 'stat_col' containing data
##how:  join tibbles together.  
##how:  aes() can reference passed element,., which is joined data.
#how: passed element contains data, and stat_col can be used to access 
#the specific element by name [[stat_col]]
##post: returns a ggplot of sf shape

cmty_stat_plot<-function(x, x_col,mainTitle="", fCol=waiver()){
     inner_join(communities, x)%>%
          ggplot()+
          geom_sf(aes(fill=.data[[x_col]]))+
          scale_fill_viridis_c(option = "C", name=fCol)+
          labs(title=mainTitle)
}

#Graphs...
png("../img/biz_ct_map.png", width = 6.5, height = 6.5, units="in", res=144)
#show businesses per community.
cmty_stat_plot(biz_by_community,"biz.ct", "Businesses per Community", "Count")
dev.off()

png("../img/biz_year_map.png", width = 6.5, height = 6.5, units="in", res=144)  
#show mean years in business for community.
cmty_stat_plot(cmty_biz_yrs,"cmty.biz.yrs", "Number of Years Inspected, 2000-2018", "Years") 
dev.off()
png("../img/biz_insp_per_yr.png", width=6.5, height = 6.5, units="in", res=144)  
#Show inspections per business per year
cmty_stat_plot(inspect_rate_by_cmty, "yr.cmty.insp.rate", "Inspections per Business per Year", "Insp. per Year")
dev.off()
#overlay of mean violations / inspection on community business count
#this is interesting.  Look at the higher number of violations in outer areas.
png("../img/biz_viol_vs_cmty.png", width = 6.5, height= 6.5, units="in", res=144)
cmty_stat_plot(cmty_viol_per_inspect,"cmty_viol_rate", "Mean Violations per Inspection", "Mean Viol.")
dev.off()

cmty_results<-inspect_long%>%
     filter(Results %in% c("Fail", "Pass w/ Conditions", "Pass"))%>%
     dplyr::select(community, Results,`Inspection ID`)%>%
     group_by(community, Results)%>%
     distinct()%>%
     dplyr::summarise(insp.ct=n())%>%
     mutate(insp.pct=insp.ct/sum(insp.ct)*100)

png("../img/insp_fail_vs_cmty.png", width = 6.5, height = 6.5, units="in", res=144)
cmty_results%>%filter(Results=="Fail")%>%
     cmty_stat_plot("insp.pct", "Percent Failed Inspections by Community", "Percent Failed")
dev.off()
```

```{r, fig.width=18, fig.height=18}
inspections.geo.f<-inspections.geo%>%
     filter(Results%in%c("Fail"), `Inspection Type` %in% c(canvass.inspection,complaint.inspection))

png("../img/biz_fail_map.png", width = 6.5, height = 6.5, units="in", res=144)
inner_join(demo.income,communities, by=c("Name"="community"))%>%
     st_as_sf()%>%
     ggplot()+
     geom_sf(aes(fill=PCI_2014.2018, geometry=geometry))+
     geom_sf(data=inspections.geo.f, aes(color=Results, geometry=geometry), size=1, alpha=0.8, fill="black")+
     scale_fill_viridis_c(option = "C", name="Per Capita Income")+
     scale_color_discrete(name="Results", type=c("steel blue"))+
     labs(title="Location of Businesses with Failing Inspections vs. Per Capita Income")
dev.off()

```
